---
title: "iBlastoids_SC_bulk_integration"
author: "jibsch"
date: "2021-01-18"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Gene Lengths for RPKM

```{r}
GL = read_tsv("~/references/human/Homo_sapiens.GRCh38.93.gene_lengths.tsv")
GL$symbol = mapIds(org.Hs.eg.db, GL$`Gene stable ID`, "ENSEMBL", column = "SYMBOL")
summary(duplicated(GL$symbol))
GL = GL[!duplicated(GL$symbol) & ! is.na(GL$symbol),]
Gl = as.data.frame(GL)
row.names(GL) = GL$symbol
```

## Load Libraries and create Pseudo Bulk

```{r}
blake = readRDS("../output/blake.Rds")
pet = readRDS("../output/seurat_pet3.Rds")
blake.TE = rowSums(as.matrix(blake@assays$RNA@counts[,blake$group == "TE"]))
blake.PE = rowSums(as.matrix(blake@assays$RNA@counts[,blake$group == "PE"]))
blake.EPI = rowSums(as.matrix(blake@assays$RNA@counts[,blake$group == "EPI"]))

iblast.TE.rep1 = rowSums(as.matrix(seurat_d6@assays$RNA@counts[,seurat_d6$cell_type2 == "TE" & seurat_d6$donor == "donor0"]))
iblast.TE.rep2 = rowSums(as.matrix(seurat_d6@assays$RNA@counts[,seurat_d6$cell_type2 == "TE" & seurat_d6$donor != "donor0"]))
iblast.PE.rep1 = rowSums(as.matrix(seurat_d6@assays$RNA@counts[,seurat_d6$cell_type2 == "PE" & seurat_d6$donor == "donor0"]))
iblast.PE.rep2 = rowSums(as.matrix(seurat_d6@assays$RNA@counts[,seurat_d6$cell_type2 == "PE" & seurat_d6$donor != "donor0"]))
iblast.EPI.rep1 = rowSums(as.matrix(seurat_d6@assays$RNA@counts[,seurat_d6$cell_type2 == "EPI" & seurat_d6$donor == "donor0"]))
iblast.EPI.rep2 = rowSums(as.matrix(seurat_d6@assays$RNA@counts[,seurat_d6$cell_type2 == "EPI" & seurat_d6$donor != "donor0"]))

pet.TE.E5 = rowSums(as.matrix(pet@assays$RNA@counts[,pet$group == "E5_TE"]))
pet.PE.E5 = rowSums(as.matrix(pet@assays$RNA@counts[,pet$group == "E5_PE"]))
pet.EPI.E5 = rowSums(as.matrix(pet@assays$RNA@counts[,pet$group == "E5_Epi"]))

pet.TE.E6 = rowSums(as.matrix(pet@assays$RNA@counts[,pet$group == "E6_TE"]))
pet.PE.E6 = rowSums(as.matrix(pet@assays$RNA@counts[,pet$group == "E6_PE"]))
pet.EPI.E6 = rowSums(as.matrix(pet@assays$RNA@counts[,pet$group == "E6_Epi"]))

pet.TE.E7 = rowSums(as.matrix(pet@assays$RNA@counts[,pet$group == "E7_TE"]))
pet.PE.E7 = rowSums(as.matrix(pet@assays$RNA@counts[,pet$group == "E7_PE"]))
pet.EPI.E7 = rowSums(as.matrix(pet@assays$RNA@counts[,pet$group == "E7_Epi"]))
```

```{r}
okae = read.table("../data/Arima_logFPKM.txt", header = TRUE, sep="\t")
okae = okae[!duplicated(okae$Gene.symbol),]
row.names(okae) = okae$Gene.symbol

okae_group = gsub("_rep[1-3]","",names(okae)[-c(1:2)])
```


## Select Okae and Linneberg Samples
```{r}
okae.select = startsWith(names(okae), "TSCT") | startsWith(names(okae), "TSblast")
cpm = removeBatchEffect(cbind(cpm(dge, log = TRUE), okae[g,okae.select]), batch = c(rep("B",3), rep("L",3), rep("P",9), rep("O",4)))


```{r}
linneberg = read.csv("../data/GSE138012_supplemental_data_1.csv", header=TRUE, as.is = TRUE)
linneberg$gene_symbol = mapIds(org.Hs.eg.db, linneberg$gene_id, "ENSEMBL", 
                               column = "SYMBOL")
summary(duplicated(linneberg$gene_symbol) & !is.na(linneberg$gene_symbol))
linneberg = linneberg[!duplicated(linneberg$gene_symbol) & !is.na(linneberg$gene_symbol),]
row.names(linneberg) = linneberg$gene_symbol

l = names(linneberg)
linneberg.select = startsWith(l, "H9_t2iLGo_PrE") | startsWith(l, "H9_t2iLGo_nEnd") | 
  #startsWith(l, "H9_MEF_pES")  | 
  startsWith(l, "H9_t2iLGo_nES")
linneberg.group = gsub("_R[1-3]","",l[linneberg.select])
```

## Integrate Data


```{r}
  

dge = DGEList(cbind(data.frame(blake.TE = blake.TE[g]/data.frame(GL)[g,3], 
                               blake.EPI = blake.EPI[g]/data.frame(GL)[g,3],
                         blake.PE = blake.PE[g]/data.frame(GL)[g,3],
                         iblast.TE.rep1 = iblast.TE.rep1[g], 
                         iblast.EPI.rep1 = iblast.EPI.rep1[g],
                         iblast.PE.rep1 = iblast.PE.rep1[g],
                         iblast.TE.rep2 = iblast.TE.rep2[g],
                         iblast.EPI.rep2 = iblast.EPI.rep2[g],
                         iblast.PE.rep2 = iblast.PE.rep2[g],
                         pet.TE.E5 = pet.TE.E5[g], pet.EPI.E5 = pet.EPI.E5[g],
                         pet.PE.E5 = pet.PE.E5[g],
                         pet.TE.E6 = pet.TE.E6[g], pet.EPI.E6 = pet.EPI.E6[g],
                         pet.PE.E6 = pet.PE.E6[g],
                         pet.TE.E7 = pet.TE.E7[g], pet.EPI.E7 = pet.EPI.E7[g],
                         pet.PE.E7 = pet.PE.E7[g]
                         ),
                    linneberg[g,linneberg.select]))

dge = calcNormFactors(dge)


cpm = removeBatchEffect(cbind(cpm(dge, log = TRUE), okae[g,okae.select]), batch = c(rep("B",3), rep("L",6), rep("P",9), rep("Lin",9), rep("O",4)))
#rep("B",3), 

pca = prcomp(t(cpm))

```


```{r}
cols = c("EPI" = "#4169E1", 
            "PE" = "#698B22", 
            "TE" = "#A020F0", 
            "H9_t2iLGo_nEnd" = "turquoise2",
            "H9_t2iLGo_nES" = "orange",
            "H9_t2iLGo_PrE" = "red",
            "TSblast" = "black", "TSCT" = "darkgrey",
            "Int_IM1" = "#8B1A1A", "Int_IM2" = "#FF3030", "Int_IM3" = "#F08080",
            "Int_nonReprog" = "black")

label = pca$x %>% as_tibble(rownames = "sample") %>%
  mutate(b=b, t=t, bt = paste(b,t,sep="_")) %>%
  group_by(bt) %>%
  summarise(PC1=mean(PC1), PC2=mean(PC2), t=t[1])



as_tibble(pca$x, rownames = "sample") %>% 
  mutate(b=b, t=t) %>%
  ggplot(aes(PC1, PC2)) + 
  geom_point(aes(colour = t), size=2) +
  #geom_text_repel(aes(colour=t, label=sample), show.legend = FALSE, size=1.5) +
  labs(colour="type", x = "PC1 (34%)", y="PC2 (18%)") +
  scale_color_manual(values = cols) +
  plotto_theme_panel() + 
  theme(legend.position = "bottom") -> p

ggsave(p, filename = "../plots3/SuppfigX_integration.pdf", width = 9, height=9, units = "cm")

p = p + geom_text_repel(data=label, aes(x=PC1, y=PC2,colour=t, label=bt), 
                  show.legend = FALSE, size=2.5, point.padding = 0.5) 

ggsave(p, filename = "../plots3/SuppfigX_integration2.pdf", width = 9, height=9, units = "cm")
```

## Soft Data

```{r}
as_tibble(pca$x, rownames="sample") %>%
  dplyr::select(sample, PC1, PC2) %>% 
  mutate(type = t) -> t

write.csv(t, row.names = FALSE, quote = FALSE, file = "../output/soft_data_table_bulk_integration_pca.csv")
```

